---
title: "Español Médico"
author: "Dom Grisafe"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(cowplot)
theme_set(theme_cowplot())
library(sna)
library(RColorBrewer)
```

# Export from Anki

GUI procedure for exporting text file of cards from Anki software

1) Anki

2) File --> Export

3) Export Window:
   - Export format: "Notes in Plain Text (*.txt)
   - Include: "Medicina"
   - Check: Include tags
   - Check: Include HTML and media references

```{r}
# load the .txt file
raw_text <- read.delim("medicina_20201115.txt", header = FALSE, sep = "\t")

# subset the column with the tag names
tag_names <- data.frame(tag_names = raw_text$V3)
```


# Data Cleaning

```{r}
# formatting variables

  # lecture factor formatting
  fct_lecture <- function(x_tag){
    case_when(
      str_detect(x_tag, pattern = "^0") ~ 1,
      str_detect(x_tag, pattern = "^1") ~ 2,
      # str_detect(x_tag, pattern = "^2") ~ 3,
      str_detect(x_tag, pattern = "^3") ~ 3,
      str_detect(x_tag, pattern = "^4") ~ 4,
      str_detect(x_tag, pattern = "^5") ~ 5,
      str_detect(x_tag, pattern = "^6") ~ 6,
      str_detect(x_tag, pattern = "^7") ~ 7
      ) %>% 
      factor(
        levels = 1:7,
        labels = c("Jerarquía", "Historial Médico", "Signos Vitales", "Nervios Craneales", "Neurológico", "Musculo-esquelético", "Cardiovascular Respiratorio")
        # , "Números y Letras"
      )
  }

  # format tag names
  prettify_tags <- function(nasty_tag){
    
    nasty_tag %>% 
      
      # remove any quantity of repeating numbers and periods in front of the tag, 
      #   then remove first underscore
      str_remove(pattern = "^[\\d+\\.+]+\\_") %>% 
      
      # replace middle underscores with spaces
      str_replace_all(pattern = "\\_", replacement = " ") %>% 
      
      # capitalize first word
      str_to_title() %>% 
      
      # remove nervios craneales roman numerals and abbreviation
      str_remove_all(pattern = "Nc |I |Ii |Iii |Iv |V |Vi |Vii |Viii |Ix |X |Xi |Xii ")

  }
  # prettify_tags("2.3_this_is_a_tag")
  # prettify_tags("2.3.3.2.5_this_is_a_tag")
  # prettify_tags("4.8_nc_IX_X_XII_músculos_paladar_lengua")
```


## Dataframes

```{r}
# each row is a card, split tags into individual columns
n_col <- 9 # max number of tags in any given row
df_cards <- tag_names %>% 
  separate(col = tag_names, sep = "\\s", into = paste0("tag_", 1:n_col), fill = "right", extra = "warn") %>% 
  dplyr::arrange(tag_names) %>% 
  rowid_to_column(var = "card_id")
    # if there are too many tags in a given row, it will warn saying:
    # "Expected n_col pieces. Additional pieces discarded in x rows [x]" 
# df_cards %>% glimpse

# each row is a tag
df_tags <- df_cards %>% 
  pivot_longer(cols = starts_with("tag_"), names_prefix = "tag_", names_to = "tag_n", values_to = "tag_name") %>% 
  # remove unnecessary tags
  dplyr::mutate(tag_name = str_replace_all(string = tag_name, pattern = "Oaxaca2019|medicina", replacement = NA_character_)) %>%
  # eliminate any rows with missing tag names
  dplyr::filter(!is.na(tag_name)) %>% 
  group_by(tag_name) %>% 
  dplyr::mutate(tag_id = cur_group_id()) %>% 
  ungroup() %>% 
  dplyr::arrange(tag_id) %>% 
  # create nodes for each deck the cards come from
  dplyr::mutate(
    lecture_id = str_extract(tag_name, pattern = "^[\\d+\\.+]+"),
    lecture_name = fct_lecture(lecture_id),
    tag_name = prettify_tags(tag_name)
  )
```

## Style Plots

```{r}
# create vector of colors
  # display.brewer.all(colorblindFriendly = TRUE) # look at all available colors
lec_name <- levels(df_tags$lecture_name)
lec_color <- brewer.pal(n = length(lec_name), "Dark2")
names(lec_color) <- lec_name

# common theme parameters for plots
set_theme_params <- function(gg_obj){
  gg_obj + 
    theme(
      legend.position = "right",
      legend.title = element_blank(),
      legend.text = element_text(size = 8)
    )
}
```

## Analysis Plan

```{r}
# count numbers of tags, cards, and lectures
n_tag     <- length(unique(df_tags$tag_name))
m_card    <- length(unique(df_tags$card_id))
k_lecture <- length(unique(df_tags$lecture_name))
```

### Summarized Lectures

We will first describe the data from the perspective of each lecture  

* Each **lecture** (k = `r k_lecture`) is a **node**
* Each **card** (m = `r m_card`) represent **edges** or links between tags
* Each **lecture** (k = `r k_lecture`) will be indicated by a different **color**

### Detailed Tags

We will then visualize the network from each tag, while indicating each lecture by color

* Each **tag** (n = `r n_tag`) is a **node**  
* Each **card** (m = `r m_card`) represent **edges** or links between tags
* Each **lecture** (k = `r k_lecture`) will be indicated by a different **color**


## Node Matrix

### Summarized Lectures

What is the number of cards if they are counted multiple times for every lecture they belong to?

```{r}
node_lec_weight <- df_tags %>% 
  group_by(card_id, lecture_name) %>% 
  summarise(card_lec_weight = n(), .groups = "drop") 
  # dplyr::mutate(lec_id = as.integer(lecture_name))
node_lec_weight %>% glimpse
```

How many cards are associated with each lecture?

```{r}
# node matrix of lectures and card weights
node_lec_cards <- node_lec_weight %>% 
  group_by(lecture_name) %>% 
  summarise(card_weight = n(), .groups = "drop")
node_lec_cards %>% glimpse
```

```{r}
# node_cards_weight %>% glimpse
node_lec_cards %>% 
  ggplot(aes(x = lecture_name, y = card_weight, fill = lecture_name)) +
  geom_col() +
  scale_fill_manual(values = as.character(lec_color)) +
  ylab(paste0("Cards (m = ", m_card, ")")) +
  xlab(paste0("Lectures (k = ", k_lecture, ")")) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 20, hjust = 1)
    ) 
```

### Detailed Tags{.tabset}

Node matrix with weights

#### Tags

How many cards are there per tag?

Note, some tags had the same name, but were from different lectures. When removing the tag ID number, these tags are associated with multiple lectures.

```{r}
# node matrix of tag id, tag names, card weights
node_tags_weight <- df_tags %>% 
  group_by(tag_id, tag_name, lecture_id) %>% 
  summarise(card_weight = n(), .groups = "drop") %>% 
  # create nodes for each deck the cards come from
  dplyr::mutate(lecture_name = fct_lecture(lecture_id))

# weight by tag, because some repeat tag names by different lecture
node_tags_weight_order <- df_tags %>% group_by(tag_name) %>% summarise(tag_weight = n(), .groups = "drop")
# node_tags_weight_order %>% glimpse

# join the tag weights for sorting plot
node_tags_weight <- node_tags_weight %>% left_join(node_tags_weight_order, by = "tag_name")
node_tags_weight %>% glimpse

# look for entries with differing card and tag weights 
#   (i.e., redundant tag names from different lectures)
# node_tags_weight %>% filter(card_weight != tag_weight) %>% arrange(-tag_weight) %>% glimpse
```

```{r}
# node_tags_weight %>% glimpse
node_tags_weight %>% 
  ggplot(aes(x = fct_reorder(tag_name, tag_weight), y = card_weight, fill = lecture_name)) +
  geom_col() +
  scale_fill_manual(values = as.character(lec_color)) +
  xlab(paste0("Tags (n = ", n_tag, ")")) +
  ylab(paste0("Cards (m = ", m_card, ")")) +
  theme(axis.text.y = element_text(size = 5, hjust = 1)) +
  coord_flip() +
  theme(legend.position = "right", legend.title = element_blank(), legend.text = element_text(size = 8))
```

#### Cards

How many tags are there per card?

```{r}
# node matrix of card id and tag weights
node_cards_weight <- df_tags %>% 
  group_by(card_id) %>% 
  summarise(tag_weight = n(), .groups = "drop")
node_cards_weight %>% glimpse
```

```{r}
# node_cards_weight %>% glimpse
node_cards_weight %>% 
  ggplot(aes(x = tag_weight)) +
  geom_histogram(binwidth = 1) +
  ylab(paste0("Cards (m = ", m_card, ")")) +
  xlab(paste0("Tags (n = ", n_tag, ")")) +
  scale_y_continuous(limits = c(0, 400), breaks = seq(0, 400, 50)) +
  coord_flip()
```


## Edge Matrix

### Summarized Lectures

```{r}
# how to get pairwise comparisons of a variable within each group
# https://intellipaat.com/community/45798/create-combination-of-all-variables-within-a-group

# create columns with pairwise comparisons
# but there is directionality and redundancy
combo_messy_lec <- df_tags %>%
  dplyr::group_by(card_id) %>% 
  dplyr::summarise(combo = list(purrr::cross_df(list(a = lecture_name, b = lecture_name))), .groups = "drop") %>% 
  tidyr::unnest(combo)
# combo_messy_lec %>% nrow

# sort rows of pairwise comparisons so can identify redundant groupings
combo_clean_lec <- combo_messy_lec %>% dplyr::select(a, b) %>% 
  t() %>% as.data.frame() %>% 
  dplyr::mutate_all(sort) %>% t() %>% 
  as.data.frame() %>% 
  bind_cols(card_id = combo_messy_lec$card_id, .)
# combo_clean_lec %>% nrow

# clean pairwise comparisons and group to get weights
edge_weights_lec <- combo_clean_lec %>% 
  # remove comparisons with redundant pairings
  dplyr::filter(V1 != V2) %>% 
  # create single variable that pairs two comparisons
  dplyr::mutate(pair_comp = paste(V1, V2, sep = "—")) %>% 
  # get weights to find repeated associations
  dplyr::group_by(pair_comp) %>% 
  dplyr::summarise(weight = n(), .groups = "drop") %>% 
  # separate pairwise comparisons into 2 variables
  tidyr::separate(col = pair_comp, into = c("from", "to"), sep = "—", remove = TRUE)
edge_weights_lec %>% glimpse
```


### Detailed Tags

```{r}
# how to get pairwise comparisons of a variable within each group
# https://intellipaat.com/community/45798/create-combination-of-all-variables-within-a-group

# create columns with pairwise comparisons
# but there is directionality and redundancy
combo_messy_tags <- df_tags %>%
  dplyr::group_by(card_id) %>%
  dplyr::summarise(combo = list(purrr::cross_df(list(a = tag_name, b = tag_name))), .groups = "drop") %>% 
  tidyr::unnest(combo)
# combo_messy_tags %>% nrow

# sort rows of pairwise comparisons so can identify redundant groupings
combo_clean_tags <- combo_messy_tags %>% dplyr::select(a, b) %>% 
  t() %>% as.data.frame() %>% 
  dplyr::mutate_all(sort) %>% t() %>% 
  as.data.frame() %>% 
  bind_cols(card_id = combo_messy_tags$card_id, .)
# combo_clean_tags %>% nrow

# clean pairwise comparisons and group to get weights
edge_weights_tag <- combo_clean_tags %>% 
  # remove comparisons with redundant pairings
  dplyr::filter(V1 != V2) %>% 
  # create single variable that pairs two comparisons
  dplyr::mutate(pair_comp = paste(V1, V2, sep = "—")) %>% 
  # get weights to find repeated associations
  dplyr::group_by(pair_comp) %>% 
  dplyr::summarise(weight = n(), .groups = "drop") %>% 
  # separate pairwise comparisons into 2 variables
  tidyr::separate(col = pair_comp, into = c("from", "to"), sep = "—", remove = TRUE)
edge_weights_tag %>% glimpse
```


# Minimal Plot{.tabset}

Notes from [Network visualization with R](https://kateto.net/sunbelt2019)

Section [5.2 A ggraph package example (for ggplot2 users)](https://kateto.net/sunbelt2019#a-ggraph-package-example-for-ggplot2-users)

Data for examples are [available here](http://www.kateto.net/wordpress/wp-content/uploads/2019/06/sunbelt2019.zip)

```{r, message=FALSE}
library(igraph)
library(ggraph)
```

## Anki Summarized Lectures

```{r}
# node matrix
# node_lec_cards %>% glimpse

# edge matrix
# edge_weights_lec %>% glimpse

net_tag <- graph_from_data_frame(directed=F,
  vertices = node_lec_cards,
  d = edge_weights_lec
  )

# Set edge width based on weight:
E(net_tag)$width <- E(net_tag)$weight

ggraph(net_tag, layout = 'linear') +
  geom_edge_arc0(aes(width = width), linetype = 1, color="gray60", strength = 0.9, alpha = 0.5) +
  geom_node_point(color = lec_color, size = 15, position = position_dodge(1)) +
  geom_node_text(aes(label = str_wrap(name, 10)), vjust = 0.5, size = 2.4) +
  theme(
    legend.position = "none",
    plot.background = element_rect(size = 10),
    plot.margin = margin(1, 1, 1, 1, "cm")
    )
```

## E.g. Media

```{r}
# load example data
nodes <- read.csv("Dataset1-Media-Example-NODES.csv", header=T, as.is=T)
links <- read.csv("Dataset1-Media-Example-EDGES.csv", header=T, as.is=T)
# head(nodes)
# head(links)

# create igraph object
net <- graph_from_data_frame(d=links, vertices=nodes, directed=T) 
# net
```

```{r}
ggraph(net, layout = 'linear') + 
    geom_edge_arc(color = "orange", width=0.7) +
    geom_node_point(size=5, color="gray50") +
    theme_void()
```


# Network Visualization

Excellent introduction to [Network Visualization with ggplot2](https://journal.r-project.org/archive/2017/RJ-2017-023/RJ-2017-023.pdf)

I found the **geomnet** package to be the most useful and intuitive, as it is most similar to ggplot2 and is least reliant on background knowledge of network analysis.

## ggnet2{.tabset}

```{r, message=FALSE}
# p.30
library(GGally)
library(network)
```

### Anki Detailed Tags

```{r}
# data step for both ggnet2 and ggnetwork
# create undirected network
tag.net <- network(edge_weights_tag[,1:2], directed = FALSE) 
tag.net # glance at network object

## Network attributes:
##  vertices = 58 
##  directed = FALSE 
##  hyper = FALSE 
##  loops = FALSE 
##  multiple = FALSE 
##  bipartite = FALSE 
##  total edges= 162 
##    missing edges= 0 
##    non-missing edges= 162 
## Vertex attribute names: 
##    vertex.names 
## No edge attributes

# assign lectures to the nodes
tag.net %v% "node_color" <- as.integer(node_tags_weight$lecture_name)

# assign weights to the edges of the network
tag.net %e% "edge_weight" <- edge_weights_tag$weight / 25

set.seed(10052016)
ggnet2(tag.net, mode = "kamadakawai", labelon = TRUE, 
       edge.size = "edge_weight",
       color =  "node_color" ,
       size = 4, 
       vjust = -0.6, label.size = 3)
```

### E.g. Madmen
```{r}
# make the data available
data(madmen, package = 'geomnet')
# data step for both ggnet2 and ggnetwork
# create undirected network
mm.net <- network(madmen$edges[, 1:2], directed = FALSE) 
mm.net # glance at network object
## Network attributes:
##   vertices = 45
##   directed = FALSE
##   hyper = FALSE
##   loops = FALSE
##   multiple = FALSE
##   bipartite = FALSE
##   total edges= 39
##     missing edges= 0
## non-missing edges= 39 ##
## Vertex attribute names:
## vertex.names
##
## No edge attributes
# create node attribute (gender) 
rownames(madmen$vertices) <- madmen$vertices$label
mm.net %v% "gender" <- as.character(
  madmen$vertices[ network.vertex.names(mm.net), "Gender"] 
)
# gender color palette
mm.col <- c("female" = "#ff69b4", "male" = "#0099ff") # create plot for ggnet2
set.seed(10052016)
ggnet2(mm.net, color = mm.col[ mm.net %v% "gender" ],
labelon = TRUE, label.color = mm.col[ mm.net %v% "gender" ], size = 2, vjust = -0.6, mode = "kamadakawai", label.size = 3)
```

## geomnet{.tabset}

```{r, message=FALSE}
# also loads ggplot2
library(geomnet)
```

### Anki Detailed Tags

```{r}
# data step: join the edge and node data with a fortify call 
TAGnet <- suppressMessages(fortify(
  as.edgedf(as.data.frame(edge_weights_tag)), 
  dplyr::select(node_tags_weight, tag_name, card_weight, lecture_name)
  ))

# layout algorithms available at
# ?gplot.layout 

# create plot
set.seed(13232123)
ggplot(data = TAGnet, 
       aes(from_id = from_id, to_id = to_id, linewidth = weight/40,
           label = str_wrap(from_id, 20) # wrap labels
           )) +
  geom_net(aes(colour = lecture_name, size = card_weight, fontsize = (card_weight/60 + 2)), layout.alg = "kamadakawai",
           labelon = TRUE, vjust = 0.5, ecolour = "grey60",
           directed = FALSE, ealpha = 0.5,
           repel = FALSE, labelcolour = "gray30"
           ) +
  scale_colour_manual(values = lec_color) +
  xlim(c(-0.05, 1.05)) +
  theme_net() %>%
  set_theme_params() +
  theme(legend.position = "none")
```


### E.g. Madmen

```{r}
# data step: join the edge and node data with a fortify call 
MMnet <- fortify(as.edgedf(madmen$edges), madmen$vertices)
# create plot
set.seed(10052016)
ggplot(data = MMnet, aes(from_id = from_id, to_id = to_id)) + geom_net(aes(colour = Gender), layout.alg = "kamadakawai",
size = 2, labelon = TRUE, vjust = -0.6, ecolour = "grey60",
directed =FALSE, fontsize = 3, ealpha = 0.5) + scale_colour_manual(values = c("#FF69B4", "#0099ff")) + xlim(c(-0.05, 1.05)) +
theme_net() +
theme(legend.position = "bottom")
```


## ggnetwork{.tabset}

```{r, message=FALSE}
library(ggnetwork)
```

### Anki Detailed Tags

```{r}
# create plot for ggnetwork. uses same data created for ggnet2 function library(ggnetwork)
set.seed(10052016)

tag.net <- network(edge_weights_tag[,1:2], directed = FALSE) 

# assign weights to the edges of the network
tag.net %v% "node_color" <- as.character(node_tags_weight$lecture_name)
tag.net %e% "edge_weight" <- edge_weights_tag$weight / 25

tag.ggnet <- ggnetwork(tag.net, layout = "kamadakawai")

ggplot(data = tag.ggnet, aes(x, y, xend = xend, yend = yend)) +
  geom_edges(color = "grey50", aes(size = edge_weight), show.legend = FALSE) + # draw edge layer 
  geom_nodes(aes(color = node_color), size = 3) + # draw node layer 
  geom_nodetext(aes(color = node_color, label = vertex.names), size = 3, vjust = -0.6) + # draw node label layer 
  # scale_colour_manual(values = tag.ggnet) +
  xlim(c(-0.05, 1.05)) + theme_blank() + theme(legend.position = "bottom") %>% 
  set_theme_params()
```


### E.g. Madmen

```{r}
# create plot for ggnetwork. uses same data created for ggnet2 function library(ggnetwork)
set.seed(10052016)
mm.ggnet <- ggnetwork(mm.net, layout = "kamadakawai")
ggplot(data = mm.ggnet,
       aes(x, y, xend = xend, yend = yend)) +
  geom_edges(color = "grey50") + # draw edge layer 
  geom_nodes(aes(colour = gender), size = 2) + # draw node layer 
  geom_nodetext(aes(colour = gender, label = vertex.names),
                size = 3, vjust = -0.6) + # draw node label layer 
  scale_colour_manual(values = mm.col) +
  xlim(c(-0.05, 1.05)) + theme_blank() + theme(legend.position = "bottom")
```




## Interactive Network Visualization{.tabset}

Notes from [Network visualization with R](https://kateto.net/sunbelt2019)

Section [6.4 Interactive JS visualization with networkD3](https://kateto.net/sunbelt2019#interactive-js-visualization-with-networkd3)

> We will also take a quick look at networkD3 which - as its name suggests - generates interactive network visualizations using the D3 javascript library. If you do not have the networkD3 library, install it with install.packages("networkD3").

```{r, message=FALSE}
library(networkD3)
```

### Anki Detailed Tags

Transform columns of edge list to numerics that start at zero.

```{r}
# links %>% glimpse
# edge_weights_tag %>% glimpse

tag_links_d3 <- data.frame(
  from = match(edge_weights_tag$from, node_tags_weight$tag_name) - 1,
  to = match(edge_weights_tag$to, node_tags_weight$tag_name) - 1,
  link_width = (edge_weights_tag$weight * 10)
  )
```

Arrange the nodes so in the same order as the “source” column in links

```{r}
# nodes %>% glimpse
# node_tags_weight %>% glimpse

tag_nodes_d3 <- cbind(
  idn = factor(node_tags_weight$tag_id, labels=node_tags_weight$tag_name), 
  node_tags_weight,
  color_lecture = str_to_lower(lec_color[match(node_tags_weight$lecture_name, names(lec_color))])
  ) %>% 
  dplyr::mutate(tag_weight = 4*tag_weight) # scale node size
```

```{r}
# color pal for getting java script this example:
# JS('d3.scaleOrdinal().domain(["1", "2", "3"]).range(["#000000", "#111111", "#222222"])')
col_node_pal <- paste0(
  "d3.scaleOrdinal().domain([\"", 
  paste0(1:7, collapse = '", "'), 
  "\"]).range([\"", 
  paste0(lec_color, collapse = '", "'), 
  "\"])"
  )

# plot interactive network
forceNetwork(
  Links = tag_links_d3, Nodes = tag_nodes_d3, Source = "from", Target = "to",
  NodeID = "idn", Group = "lecture_name", colourScale = JS(col_node_pal), # linkWidth = "link_width", 
  linkColour = "#afafaf", fontSize = 20, zoom = F, legend = F,
  Nodesize = 7, opacity = 0.8, charge = -300, 
  width = 1000, height = 1000
  )
```

### E.g. Media

> The data that this library needs from is in the standard edge list form, with a few little twists. In order for things to work, the node IDs have to be numeric, and they also have to start from 0. An easy was to get there is to transform our character IDs to a factor variable, transform that to numeric, and make sure it starts from zero by subtracting 1.

```{r}
links.d3 <- data.frame(from=as.numeric(factor(links$from))-1, 
                       to=as.numeric(factor(links$to))-1 )
```

> The nodes need to be in the same order as the “source” column in links:

```{r}
nodes.d3 <- cbind(idn=factor(nodes$media, levels=nodes$media), nodes) 
```

> Now we can generate the interactive chart. The Group parameter in it is used to color the nodes. Nodesize is not (as one might think) the size of the node, but the number of the column in the node data that should be used for sizing. The charge parameter controls node repulsion (if negative) or attraction (if positive).

```{r}
forceNetwork(Links = links.d3, Nodes = nodes.d3, Source="from", Target="to",
               NodeID = "idn", Group = "type.label", linkWidth = 1,
               linkColour = "#afafaf", fontSize=12, zoom=T, legend=T,
               Nodesize=6, opacity = 0.8, charge=-300, 
               width = 600, height = 400)
```

