---
title: "Español Médico"
author: "Dom Grisafe"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(cowplot)
theme_set(theme_cowplot())
```

# Export from Anki

GUI procedure for exporting text file of cards from Anki software

1) Anki

2) File --> Export

3) Export Window:
   - Export format: "Notes in Plain Text (*.txt)
   - Include: "Medicina"
   - Check: Include tags
   - Check: Include HTML and media references

```{r}
# load the .txt file
raw_text <- read.delim("medicina.txt", header = FALSE, sep = "\t")

# subset the column with the tag names
tag_names <- data.frame(tag_names = raw_text$V3)
```


# Clean Data

```{r}
# each row is a card, split tags into individual columns
n_col <- 9 # max number of tags in any given row
df_cards <- tag_names %>% 
  separate(col = tag_names, sep = "\\s", into = paste0("tag_", 1:n_col), fill = "right", extra = "warn") %>% 
  dplyr::arrange(tag_names) %>% 
  rowid_to_column(var = "card_id")
    # if there are too many tags in a given row, it will warn saying:
    # "Expected n_col pieces. Additional pieces discarded in x rows [x]" 
# df_cards %>% glimpse

# each row is a tag
df_tags <- df_cards %>% 
  pivot_longer(cols = starts_with("tag_"), names_prefix = "tag_", names_to = "tag_n", values_to = "tag_name") %>% 
  # remove unnecessary tags
  dplyr::mutate(tag_name = str_replace_all(string = tag_name, pattern = "Oaxaca2019|medicina", replacement = NA_character_)) %>%
  # eliminate any rows with missing tag names
  dplyr::filter(!is.na(tag_name)) %>% 
  group_by(tag_name) %>% 
  dplyr::mutate(tag_id = cur_group_id()) %>% 
  ungroup() %>% 
  dplyr::arrange(tag_id) %>% 
  # create nodes for each deck the cards come from
  dplyr::mutate(
    lecture = case_when(
      str_detect(tag_name, pattern = "^0") ~ 1,
      str_detect(tag_name, pattern = "^1") ~ 2,
      str_detect(tag_name, pattern = "^3") ~ 3,
      str_detect(tag_name, pattern = "^4") ~ 4,
      str_detect(tag_name, pattern = "^5") ~ 5,
      str_detect(tag_name, pattern = "^6") ~ 6
    ) %>% factor(levels = 1:6, labels = c("Jerarquía", "Historial Médico", "Signos Vitales", "Nervios Craneales", "Examen Neurológico", "Examen Musculoesquelético"))
  )
```

The following will be used to visualize the network  

* Each **tag** (n = `r length(unique(df_tags$tag_name))`) is a **node**  
* Each **card** (n = `r length(unique(df_tags$card_id))`) represent **edges** or links between tags
* Each **lecture** (n = `r length(unique(df_tags$lecture))`) will be indicated by a different **color**


# Node Graph

```{r}
# show number of tags identified within each lecture
df_tags %>% 
  ggplot() +
  geom_bar(aes(x = lecture)) +
  scale_y_continuous(limits = c(0,300)) +
  theme(
    axis.text.x = element_text(angle = 20, hjust = 1)
  )
```


# Routes Graph

## Edge List

```{r}

```


```{r}
# library(igraph)
# routes_igraph <- graph_from_data_frame(d = edges, vertices = nodes, directed = TRUE)
# routes_igraph_tidy <- as_tbl_graph(routes_igraph)
# 
# ggraph(routes_igraph, layout = "linear") + 
#   geom_edge_arc(aes(width = weight), alpha = 0.8) + 
#   scale_edge_width(range = c(0.2, 2)) +
#   geom_node_text(aes(label = label)) +
#   labs(edge_width = "Letters") +
#   theme_graph()
```

# network

```{r}
# library(network)
# nw_tags <- network(df_tags, )
# plot(nw_tags)
```



# 7.1 Interactive plots of static networks

We will start from plots of static network portraying a single time point (or an aggregated time period).


As ndtv is part of the Statnet family, it will accept objects from the network package such as the one we created earlier (net3).

```{r}
# # install.packages('ndtv', dependencies=T)
# library('ndtv')
# net3 
```


Most of the parameters below are self-explanatory at this point (bg is the background color of the plot). Two new parameters we haven’t used before are vertex.tooltip and edge.tooltip. Those contain the information that we can see when moving the mouse cursor over network elements. Note that the tooltip parameters accepts html tags – for example we will use the line break tag <br>. The parameter launchBrowser instructs R to open the resulting visualization file (filename) in the browser.


```{r}
# par(mar=c(0,0,0,0))
# 
# render.d3movie(net3, usearrows = F, displaylabels = F, bg="#111111", 
#        vertex.border="#ffffff", vertex.col =  net3 %v% "col",
#        vertex.cex = (net3 %v% "audience.size")/8, 
#        edge.lwd = (net3 %e% "weight")/3, edge.col = '#55555599',
#        vertex.tooltip = paste("<b>Name:</b>", (net3 %v% 'media') , "<br>",
#                               "<b>Type:</b>", (net3 %v% 'type.label')),
#        edge.tooltip = paste("<b>Edge type:</b>", (net3 %e% 'type'), "<br>", 
#                             "<b>Edge weight:</b>", (net3 %e% "weight" ) ),
#        launchBrowser=F, filename="Media-Network.html" ) 
```

